### 2019-01-07 writing/thesis/chapters/methods.tex
More work on methods chapter.

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\NewDocumentCommand{\sbasis}{ O{} O{} }{\phi^{#1}_{#2}}
\NewDocumentCommand{\stbasis}{ O{} O{} }{ \Theta^{#1}_{#2}}
\NewDocumentCommand{\sbasisRef}{ O{} O{} }{\hat{\sbasis[#1][#2]}}
\NewDocumentCommand{\stbasisRef}{ O{} O{} }{\hat{\stbasis[#1][#2]}}
%\newcommand{\stestfunction}[1]{\sbasis[#1]}
\NewDocumentCommand{\stestfunction}{ O{} O{} }{\sbasis[#1][#2]}
\NewDocumentCommand{\sttestfunction}{ O{} O{} }{\stbasis[#1][#2]}
\newcommand{\normal}{\bm{n}}
\newcommand{\dsol}[1][h]{\bm{u}_{#1}}
\newcommand{\stpredictor}[1][h]{\bm{q}_{#1}}

@@

% Integrals
\newcommand{\intdt}[1]{\int_{t^n}^{t^{n+1}} #1 \dd{t}}
\newcommand{\intdcell}[1]{\int_{\cell} #1 \dd{\bm{x}}}
\newcommand{\intdrefcell}[1]{\int_{\cell} #1 \dd{\hat{\bm{x}}}}
\newcommand{\intdcellb}[1]{\int_{\partial{} \cell} #1 \dd{S}} % TODO: define boundary of cell
\newcommand{\intdrefcellb}[1]{\int_{\partial{} \cell} #1 \hat{\dd{S}}} % TODO: define boundary of cell

% Yet to be sorted
\newcommand{\quadWeight}[1][i]{w_{#1}}

@@

\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
Some of our scenarios can be described as a pertubation of a background state that is in hydrostatic balance
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
where $\backgroundPressure$ and $\backgroundRho$ is the background pressure and density.
We now focus on the following part of the momentum equation (neglecting viscous effects) of \cref{eq:equation-set}
\begin{equation}
  \label{eq:momentum-equation}
  \pdv{\Qj}{t}+ \divergence{
    \Qv \cdot \left( \bm{I} \QE + \bm{I} \pressure \right)
  }
  =
  -g \bm{k} \cdot \Qrho
\end{equation}.
The background state \cref{eq:hydrostatic-balance} leads to large values and can diminish other parts of the flux due to numerical instabilities~\cite{muller2010adaptive}.
We can now split the pressure as $\pressure = \backgroundPressure(z) + \pressure' (\bm{x}, t)$ and the density as $\Qrho = \backgroundRho(z) + \Qrho'(\bm{x}, t)$.
Inserting this splitting, the definition of divergence into \cref{eq:momentum-equation} allows us to apply \cref{eq:hydrostatic-balance}.
Furthermore, the background states have zero derivatives in all directions but in $z$.
We arrive at 
\begin{equation}
  \label{eq:momentum-equation}
  \pdv{\Qj}{t}+ \divergence \left(
    \Qv \cdot \left( \bm{I} \QE + \bm{I} \pressure' \right)
   \right)
  =
  -g \bm{k} \Qrho'.
\end{equation}.
We thus cancel parts of the flux with parts of the source term.
Note that the time derivative of the background density also vanishes.
We do not use this to simplify the implementation of the gradient computation.
This form of the equation is similar to equation set 3 of~\cite{giraldo2008study} but without splitting the energy.

@@

\sidetitle{Cauchy}
For some scenarios we use Cauchy boundary conditions.
In most cases, we would like to impose periodic boundary conditions, due to the inner workings of \exahype{} this is not possible.
Instead we use the analytical solution of our problems at the boundary, imposing both value and gradients of the conservative variables.
Note that this leads to an error when our problem does not posses an exact analytical solution.
This is the case for test cases that are analytical solutions to the incompressible Navier Stokes equations but do not satisfy the compressible equation set.

@@

We manually set the energy component of the numerical flux to zero for both no- and free-slip conditions.
\section{The ADER-DG Method}\label{sec:ader-dg}
We describe the arbitrary derivative discontinous Galerkin (\aderdg) method in this chapter.
Our description of the main method follows~\cite{dumbser2008unified,dumbser2010arbitrary,dumbser2018efficient}, our notation follows primarily~\cite{dumbser2018efficient}.
We discuss the approximation of systems of partial differential equations (\pde) of the form
\begin{equation}
  \label{eq:conservation-law-gradient}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q, \gradQ) = \source(\bm{x}, t, \Q),
\end{equation}
which in contrast to hyperbolic conservation laws of the form
\begin{equation}
  \label{eq:conservation-law}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q) = \source(\bm{x}, t, \Q)
\end{equation}
contain the gradient $\gradQ$ of the so called vector of conservative variables $\Q$~\cite{dumbser2010arbitrary}.
There are therefore not hyperbolic but rather parabolic or elliptic.

We discuss the solution of a hyperbolic conservation law \cref{eq:conservation-law} with domain $\domain$ and boundary $\boundary$.
Let $\domain$ and $\boundary$ denote our two- or three-dimensional computational domain and its boundary.
In our implementation of the discontinous Galerkin (\dg) framework, we approximate this solution in the space
\begin{equation}
  \label{eq:dg-space}
  \broken = \bigcup_i \cell
\end{equation}
of disjoint quadrilateral cells $\cell$.
Note that we do not distinguish between the broken approximation space and the domain, the use should be clear given its surrounding context.
In the following we make use of the Einstein summation convention where summation over repeated indexes is implied.

% Inside each cell $\cell$ we represent the solution in terms of the basis function 
% \begin{equation}
%   \label{eq:cell-approx}
%   u(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \sbasis{l}(\bm{x}),
% \end{equation}
% where $l$ is a multi-index, containing one index per spatial dimension.
% \todo{dofs $\hat{u}$, local solution, etc.}
% For example, $(l = (l_1, l_2))$ for the two dimensional case.
% This polynomial is interpolating, i.e.\ \ldots\todo{interpolating?}.
% This choice of basis (called \textit{nodal} basis) allows us to easily compute integrals over cells using Gaussian quadrature.
% In detail, we use a the Lagrange interpolation polynomials.

We now describe the derivation of the \textsc{ader-dg} method.
This scheme is a predictor-corrector method.
We first compute a local time evolution of each cell in a predictor step and then add the influence of the neighbors in the corrector step.
$\ldots$

% \begin{algorithm}[H]
%   \begin{algorithmic}
%     \Let{$\stpredictor$}{Initial guess}
%     \Let{$F_h$}{$\flux(\Q, \gradQ)$}
%     \Let{$\stpredictor$}{Solve \cref{eq:space-time-predictor} using $F_h, \Q, \gradQ$}
%     \Let{$\left( \hat{F_h}, \hat{\Q}, \hat{\gradQ}, \hat{\stpredictor} \right)$}
%         {Extrapolate to boundary using \cref{eq:boundary-extrapolation}}
%     \Let{update}{Use \cref{eq:corrector} with boundary extrapolated values (fluxes of neighbors)}
%   \end{algorithmic}
%   \caption{The \textsc{ader-dg} algorithm for one cell.}
% \end{algorithm}

\sidetitle{Reference Cell}
\newcommand{\mapping}{\mathcal{M}}
\newcommand{\volume}{V}
For reasons of computational efficiency, we describe the scheme in terms of a reference cell.
We use the quad cell $\cell = [0, 1]^D$\todo{check if $[0,1]$ detontes correct interval}.
The mapping, which takes a point on the reference cell with coordinates $\hat{\bm{x}}$ and returns a point $\bm{x}$ on a cell with center and widths $(\Delta x, \Delta y, \Delta z)$, is given by
\begin{equation}\label{eq:space-mapping}
  \hat{x} = \mapping (\bm{x}) =
  \operatorname{cell-center}(\cell) +
\begin{pmatrix}
\Delta x && \\
&\Delta y & \\
&&\Delta z&
\end{pmatrix}
 % \operatorname{diag}(\Delta x, \Delta y, \Delta z)
  \begin{pmatrix}
    x - 0.5\\
    y - 0.5\\
    z - 0.5
  \end{pmatrix}.
\end{equation}

We also define the inverse mapping $\mapping^{-1}$.
This can be used to define a function $f(\bm{x})$ acting on a point $\bm{x}$ of a regular cell in term of a function $\hat{f}(\bm{x})$ that acts on a point of the reference cell by considering
\begin{equation}
  \label{eq:function-ref-cell}
  f(\bm{x}) = f\left( \mapping^{-1}(\bm{x}) \right) = \hat{f}(\hat{\bm{x}}).
\end{equation}

Similar to this, we define a mapping that maps a point in time to the reference time $\in (0,1)$
\todo{Fix this/remove}
\newcommand{\timeMapping}{\mathcal{T}}
\begin{equation}
  \label{eq:time-mapping}
  \hat{t} = \timeMapping(t) = t_0 + (\Delta t) t,
\end{equation}
where $\Delta t$ is timestep of a cell and $t_0$ is the beginning of its time interval.

The Jacobian determinant of $\mapping$ is simply the volume $V$ of the cell
\begin{equation}
  \label{eq:determinant-mapping}
  \volume = \operatorname{det}(\operatorname{Jacobian}\mapping) = \Delta x \, \Delta y \,\Delta z,
\end{equation}
the Jacobian determinant of $\timeMapping$ is $\Delta t$.
This is useful for evaluating derivatives by the chain rule and, similarly, integrals by substitution.
For example, we can evaluate a volume integral by
\begin{equation}
  \label{eq:integration-by-substitution}
  \intdcell{
f(\bm{x})
  }
  =
\volume \intdrefcell{
    \hat{f}(\hat{\bm{x}})
  }
\end{equation}
\sidetitle{Basis Functions}
\newcommand{\lagrangeRef}[1][i]{\hat{L}_{#1}}
\newcommand{\lagrange}[1][i]{L_{#1}}
The next step is the definition of basis functions for the reference cell.
We use Langrange polynomials that are defined in the one-dimensional case by
\begin{equation}
  \label{eq:lagrange-basis}
  \lagrangeRef(x) = \prod_{i \neq j}^N \frac{x - x_j}{x_i - x_j}.
\end{equation}
We can then represent functions as a sum over the support points $x_j$
\begin{equation}
  \label{eq:langrange-expansion-1d}
  f(x) = \sum_{i = 0}^N f(x_i) \lagrangeRef[i](x) = f(x_i) \lagrangeRef[i](x),
\end{equation}
where we used the Einstein summation convention where summation over repeated indices is implied.
The support points $x_i$ are chosen such that they coincide with the nodes of the Gauss-Legendre quadrature which we can use to evaluate integrals
\begin{equation}
  \label{eq:quadrature-1d}
  \int_0^1 f(x) \dd{x} = \sum_n^N \quadWeight[n] f(x_n),
\end{equation}
where $x_n$ is a quadrature point with weight $\quadWeight[n]$.
The basis functions are similarly defined in $d$ dimensions by considering nodes that are the tensor-products of $d$ one-dimensional Gauss-Legendre quadrature.
To simplify notation, we order the polynomials by a linearised index, a $d$-dimensional function is then approximated by
\begin{equation}
  \label{eq:lagrange-expansion}
  f(\bm{x}) = \sum_{i = 0}^{N^d} f(\bm{x}_i) \lagrangeRef[i](\bm{x}) = f(\bm{x}_i) \lagrangeRef[i](\bm{x}).
\end{equation}
Similarly to the one-dimensional case, we can evaluate integrals by
\begin{align}
  \label{eq:quadrature}
  \underbrace{\int_0^1 \cdots \int_0^1}_{D \text{ times}} f(\bm{x}) \dd{\bm{x}} &= \sum_n^{(N+1)^D - 1} \quadWeight[n] f(\bm{x}_n) \\
  \intertext{with}
  \quadWeight[n] &= \prod_{i}^{d} \quadWeight[i].
\end{align}
\todo{Fix notation, especially multi-dimensional integral.}
\todo{Check number of total polynomials.}
% TODO: Quadrature

Two important properties of this family of polynomials are
  \begin{alignat}{10}
    \label{eq:basis-interpolating}
&& \text{Interpolating:} \qquad & \lagrangeRef[i] (x_j) &= \delta_{ij}\\
\label{eq:basis-orthogonality}
&& \text{Orthogonality:} \qquad &
\int_0^1 \cdots \int_0^1 \lagrangeRef[j](\hat{\bm(x)} \lagrangeRef[j]{\hat{\bm{x}}}) \dd{\hat{\bm{x}}}  &= \quadWeight[i] \delta_{ij}
    \end{alignat}

We will later need basis functions over a space-cell 
\begin{align}\label{eq:space-basis}
  \begin{split}
  \sbasisRef(\hat{\bm{x}}) &= \sum_i^{(N+1)^D - 1} \lagrange[i] (x_i)\\
  \sbasis(\bm{x}) &= \sbasisRef \left( \mapping(\bm{x}) \right)
  \end{split}
\end{align}
where $\sbasisRef$ denotes the basis function on the reference cell.
Similarly, we define a basis for a space-time element defined by $(0, 1)^D \times (t^n, t^{n+1})$
\begin{align}\label{eq:space-time-basis}
  \begin{split}
  \stbasisRef(\hat{\bm{x}}, \hat{t}) &= \sum_j^{N} \sum_i^{(N+1)^D - 1} \lagrange[j](x_j) \lagrange[i] (x_i)\\
  \stbasis (\bm{x}, t) &= \stbasisRef \left( \mapping(\bm{x}), \timeMapping(t) \right)
  \end{split}
\end{align}

\sidetitle{Predictor}
To derive the predictor, we take an approximation of our solution in a nodal basis that is defined both in space and time
\begin{equation}
  \label{eq:cell-approx-space-time}
  q(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \stbasis[][l](\bm{x}, t).
\end{equation}
Furthermore we assume that all quantities are given in this represantation.
We now multiply the conservation law by a test function (of the same function space as the basis) and arrive at the weak formulation
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t)
    \left(
      \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
=
\intdt{\intdcell{
  \sttestfunction{k} \source(\stpredictor)
}}.
\end{equation}
We integrate the first term by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$.
This corresponds to upwinding in time~\cite{dumbser2008unified}.
Note that this neglects the interaction with neighbouring cells.
We account for this in the corrector step.
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \sttestfunction{k} (\bm{x}, t^{n+1}) \stpredictor(\bm{x}, t^{n+1})
}
&-
\intdt{\intdcell{
    \pdv{}{t} \sttestfunction{k}(\bm{x}, t) \stpredictor(\bm{x}, t)
}}
= \\
\intdcell{
  \sttestfunction{k}(\bm{x},t^n) \dsol(\bm{x}, t^n)
}
&+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x},t) \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
}}
+
\intdt{\intdcell{
    \sttestfunction{k}(\bm{x}, t) \source(\stpredictor)
}}
\end{split}
\end{align}
Inserting \cref{eq:cell-approx-space-time} results in a local systems of equations that can be by a fixed point iteration scheme.
\todo{Inital guess correct? Maybe cite sth?}
As initial value for the space-time-predictor we use the solution of the timestep before.
\todo{Collect into matrices and show scheme?}
To show the computational efficiency of our scheme, we follow the approach of~\cite{dumbser2008unified} and collect our integrals into matrices\todo{Source is Dominics \aderdg-document.}.

>>>>

% Stuff for numerics
\newcommand{\domain}{\Omega}
\newcommand{\broken}{\domain}
\newcommand{\cell}[1][]{C_{#1}}
\newcommand{\boundary}{\partial \domain}
\NewDocumentCommand{\sbasis}{ O{} O{} }{\phi^{#1}_{#2}}
\NewDocumentCommand{\stbasis}{ O{} O{} }{ \Theta^{#1}_{#2}}
\NewDocumentCommand{\sbasisRef}{ O{} O{} }{\hat{\phi}^{#1}_{#2}} %{\hat{\sbasis[#1][#2]}}
\NewDocumentCommand{\stbasisRef}{ O{} O{} }{\hat{\Theta}^{#1}_{#2}} %{\hat{\stbasis[#1][#2]}}
%\newcommand{\stestfunction}[1]{\sbasis[#1]}
\NewDocumentCommand{\stestfunction}{ O{} O{} }{\sbasis[#1][#2]}
\NewDocumentCommand{\sttestfunction}{ O{} O{} }{\stbasis[#1][#2]}
\newcommand{\normal}{\bm{n}}
\newcommand{\dsol}[1][h]{\bm{u}_{#1}}
\NewDocumentCommand{\stpredictor}{ O{} O{} }{\overline{\bm{q}}^{#1}_{#2}}
\NewDocumentCommand{\spredictor}{ O{} O{} }{\bm{q}^{#1}_{#2}}

@@

% Integrals
\newcommand{\curTime}{t^n}
\newcommand{\nextTime}{t^{n+1}}
\newcommand{\intdt}[1]{\int_{\curTime}^{\nextTime} #1 \dd{t}}
\newcommand{\intdcell}[1]{\int_{\cell} #1 \dd{\bm{x}}}
\newcommand{\intdrefcell}[1]{\int_{\cell} #1 \dd{\hat{\bm{x}}}}
\newcommand{\intdcellb}[1]{\int_{\partial{} \cell} #1 \dd{S}} % TODO: define boundary of cell
\newcommand{\intdrefcellb}[1]{\int_{\partial{} \cell} #1 \hat{\dd{S}}} % TODO: define boundary of cell

% Yet to be sorted
\newcommand{\quadWeight}[1][i]{w_{#1}}
\newcommand{\quadNode}[1][i]{n_{#1}}
\newcommand{\sbasisSize}{(N+1)^\dimensions}
\newcommand{\stbasisSize}{(N+1)^{\dimensions+1}}

@@

\newcommand{\backgroundPressure}{\overline{\pressure}}
\newcommand{\backgroundRho}{\overline{\Qrho}}
Some of our scenarios can be described as a pertubation of a background state that is in hydrostatic balance
\begin{equation}
  \label{eq:hydrostatic-balance}
  \pdv{}{z} \backgroundPressure{\left (z \right )} = -g \backgroundRho(z),
\end{equation}
where $\backgroundPressure$ and $\backgroundRho$ is the background pressure and density.
We now focus on the following part of the momentum equation (neglecting viscous effects) of \cref{eq:equation-set}
\begin{equation}
  \label{eq:momentum-equation}
  \pdv{\Qj}{t}+ \divergence{
    \Qv \cdot \left( \bm{I} \QE + \bm{I} \pressure \right)
  }
  =
  -g \bm{k} \cdot \Qrho
\end{equation}.
The background state \cref{eq:hydrostatic-balance} leads to large values and can diminish other parts of the flux due to numerical instabilities~\cite{muller2010adaptive}.
We can now split the pressure as $\pressure = \backgroundPressure(z) + \pressure' (\bm{x}, t)$ and the density as $\Qrho = \backgroundRho(z) + \Qrho'(\bm{x}, t)$.
Inserting this splitting, the definition of divergence into \cref{eq:momentum-equation} allows us to apply \cref{eq:hydrostatic-balance}.
Furthermore, the background states have zero derivatives in all directions but in $z$.
We arrive at 
\begin{equation}
  \label{eq:momentum-equation-split}
  \pdv{\Qj}{t}+ \divergence \left(
    \Qv \cdot \left( \bm{I} \QE + \bm{I} \pressure' \right)
   \right)
  =
  -g \bm{k} \Qrho'.
\end{equation}.
We thus cancel parts of the flux with parts of the source term.
Note that the time derivative of the background density also vanishes.
We do not use this to simplify the implementation of the gradient computation.
This form of the equation is similar to equation set 3 of~\cite{giraldo2008study} but without splitting the energy.

@@

\sidetitle{Cauchy}
For some scenarios we use Cauchy boundary conditions, i.e.\ we prescribe both $\Q$ and $\gradQ$ at the boundary.
In most cases, we would like to impose periodic boundary conditions, due to the inner workings of \exahype{} this is not possible.
Instead we use the analytical solution of our problems at the boundary, imposing both value and gradients of the conservative variables.
Note that this leads to an error when our problem does not posses an exact analytical solution.
This is the case for test cases that are analytical solutions to the incompressible Navier Stokes equations but do not satisfy the compressible equation set.

@@

We manually set the energy component of the numerical flux to zero for both no- and free-slip conditions.
\section{The ADER-DG Method}\label{sec:ader-dg}
\newcommand{\NVar}{N^{\text{var}}}
\newcommand{\dimensions}{d}
\newcommand{\mapping}{\mathcal{M}}
\newcommand{\volume}{V}
\newcommand{\cellCenter}{\operatorname{cell-center}}
We describe the arbitrary derivative discontinous Galerkin (\aderdg) method in this chapter.
Our description of the main method follows~\cite{dumbser2008unified,dumbser2010arbitrary,dumbser2018efficient}, our notation follows primarily~\cite{dumbser2018efficient}.
Let $\NVar$ be the number of variables and $\dimensions$ the number of spatial dimensions.
We discuss the approximation of systems of partial differential equations (\pde) of the form
\begin{equation}
  \label{eq:conservation-law-gradient}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q, \gradQ) = \source(\bm{x}, t, \Q),
\end{equation}
which in contrast to hyperbolic conservation laws of the form
\begin{equation}
  \label{eq:conservation-law}
 \frac{\partial}{\partial_t}  \Q + \div \flux(\Q) = \source(\bm{x}, t, \Q)
\end{equation}
contain the gradient $\gradQ \in \mathbb{R}^{\NVar \times \dimensions}$ of the so called vector of conservative variables $\Q \in \mathbb{R}^{\NVar}$~\cite{dumbser2010arbitrary}.
They are therefore not hyperbolic but rather parabolic or elliptic.

We discuss the solution of a conservation law \cref{eq:conservation-law-gradient} with two or three dimensional domain $\domain$ and boundary $\boundary$.
In our implementation of the discontinous Galerkin (\dg) framework, we approximate this solution in the space
\begin{equation}
  \label{eq:dg-space}
  \broken = \bigcup_k \cell[k]
\end{equation}
of disjoint quadrilateral cells $\cell$.
Note that we do not distinguish between the broken approximation space and the domain, the use should be clear given its surrounding context.
For each cell we denote its center by $\cellCenter{\cell}$ and its size by $\Delta x, \Delta y$ and $\Delta z$.
%In the following we make use of the Einstein summation convention where summation over repeated indexes is implied.
% Inside each cell $\cell$ we represent the solution in terms of the basis function 
% \begin{equation}
%   \label{eq:cell-approx}
%   u(\bm{x}, t^n)_{|\cell[i]} = \hat{u}^n_{i,l} \sbasis{l}(\bm{x}),
% \end{equation}
% where $l$ is a multi-index, containing one index per spatial dimension.
% \todo{dofs $\hat{u}$, local solution, etc.}
% For example, $(l = (l_1, l_2))$ for the two dimensional case.
% This polynomial is interpolating, i.e.\ \ldots\todo{interpolating?}.
% This choice of basis (called \textit{nodal} basis) allows us to easily compute integrals over cells using Gaussian quadrature.
% In detail, we use a the Lagrange interpolation polynomials.

The \aderdg{}-scheme is a predictor-corrector method.
We first compute a local time evolution of each cell in a predictor step and then add the influence of the neighbors in the corrector step.
% \begin{algorithm}[H]
%   \begin{algorithmic}
%     \Let{$\stpredictor$}{Initial guess}
%     \Let{$F_h$}{$\flux(\Q, \gradQ)$}
%     \Let{$\stpredictor$}{Solve \cref{eq:space-time-predictor} using $F_h, \Q, \gradQ$}
%     \Let{$\left( \hat{F_h}, \hat{\Q}, \hat{\gradQ}, \hat{\stpredictor} \right)$}
%         {Extrapolate to boundary using \cref{eq:boundary-extrapolation}}
%     \Let{update}{Use \cref{eq:corrector} with boundary extrapolated values (fluxes of neighbors)}
%   \end{algorithmic}
%   \caption{The \textsc{ader-dg} algorithm for one cell.}
% \end{algorithm}

\sidetitle{Reference Cell}
For reasons of computational efficiency, we describe the scheme in terms of a reference cell.
We use the quad cell $\hat{\cell} = [0, 1]^\dimensions$ as space reference cell.\todo{check if $[0,1]$ detontes correct interval}.
The mapping, which takes a point on a grid cell with coordinates $\bm{x}$ and returns the corresponding point $\hat{\bm{x}}$ on the reference cell
\begin{equation}\label{eq:space-mapping}
  \hat{x} = \mapping (\bm{x}) =
  \cellCenter(\cell) +
\begin{pmatrix}
\Delta x && \\
&\Delta y & \\
&&\Delta z&
\end{pmatrix}
  \begin{pmatrix}
    x - 0.5\\
    y - 0.5\\
    z - 0.5
  \end{pmatrix}.
\end{equation}
We also define the inverse mapping $\mapping^{-1}$.
This can be used to define a function $f(\bm{x})$ acting on a point $\bm{x}$ of a regular cell in terms of a function $\hat{f}(\bm{x})$ that acts on the corresponding point of the reference cell by the relation
\begin{equation}
  \label{eq:function-ref-cell}
  f(\bm{x}) = f\left( \mapping^{-1}(\bm{x}) \right) = \hat{f}(\hat{\bm{x}}).
\end{equation}

Similar to this, we define a mapping that maps a point in time to the reference time $\hat{t} \in (0,1)$
\newcommand{\timeMapping}{\mathcal{T}}
\begin{equation}
  \label{eq:time-mapping}
  \hat{t} = \timeMapping(t) = t_0 + (\Delta t) t,
\end{equation}
where $\Delta t$ is timestep of a cell and $t_0$ is the beginning of its time interval.
The space-time reference cell is then $\hat{\cell} = [0,1]^{D+1}$

The Jacobian determinant of $\mapping$ is simply the volume $V$ of the cell
\begin{equation}
  \label{eq:determinant-mapping}
  \volume = \operatorname{det}(\operatorname{Jacobian}\mapping) = \Delta x \, \Delta y \,\Delta z,
\end{equation}
the Jacobian determinant of $\timeMapping$ is $\Delta t$.
This is useful for evaluating derivatives by the chain rule and, similarly, integrals by substitution.
For example, we can evaluate a volume integral by
\begin{equation}
  \label{eq:integration-by-substitution}
  \intdcell{
f(\bm{x})
  }
  =
\volume \intdrefcell{
    \hat{f}(\hat{\bm{x}})
  }
\end{equation}
\sidetitle{Basis Functions}
\newcommand{\lagrange}[1][i]{\varphi_{#1}}
\newcommand{\lagrangeRef}[1][i]{\hat{\varphi}_{#1}}
The next step is the definition of basis functions for the reference cell.
We use Langrange polynomials that are defined in the one-dimensional case by
\begin{equation}
  \label{eq:lagrange-basis}
  \lagrangeRef(x) = \prod_{i \neq j}^N \frac{x - x_j}{x_i - x_j}.
\end{equation}
We can then represent functions as a sum over the support points $x_j$
\begin{equation}
  \label{eq:langrange-expansion-1d}
  f(x) = \sum_{i = 0}^N f(x_i) \lagrangeRef[i](x) = f(x_i) \lagrangeRef[i](x),
\end{equation}
where we used the Einstein summation convention where summation over repeated indices is implied\todo{Remove this sentence, or sentence before introducing einstein summation!}.
The support points $x_i$ are chosen such that they coincide with the nodes of the Gauss-Legendre quadrature which we can use to evaluate integrals
\begin{equation}
  \label{eq:quadrature-1d}
  \int_0^1 f(x) \dd{x} = \sum_n^N \quadWeight[n] f(x_n),
\end{equation}
where $x_n$ is a quadrature point with weight $\quadWeight[n]$.

The basis functions are similarly defined in $d$-dimensions by considering nodes that are the tensor-products of $d$ one-dimensional Gauss-Legendre quadrature.
To simplify notation, we order the polynomials by a linearised index, a $d$-dimensional function is then approximated by
\begin{equation}
  \label{eq:lagrange-expansion}
  f(\bm{x}) = \sum_{i = 0}^{(N+1)^\dimensions - 1} f(\bm{x}_i) \lagrangeRef[i](\bm{x}) = f(\bm{x}_i) \lagrangeRef[i](\bm{x}).
\end{equation}
Similarly to the one-dimensional case, we can evaluate integrals by
\begin{align}
  \label{eq:quadrature}
  \underbrace{\int_0^1 \cdots \int_0^1}_{\dimensions \text{ times}} f(\bm{x}) \dd{\bm{x}} &= \sum_n^{(N+1)^D - 1} \quadWeight[n] f(\bm{x}_n) \\
  \intertext{with}
  \quadWeight[n] &= \prod_{i}^{d} \quadWeight[i].
\end{align}
\todo{Fix notation, especially multi-dimensional integral.}
\todo{Check number of total polynomials.}
% TODO: Quadrature

Two important properties of this family of polynomials are
  \begin{alignat}{10}
    \label{eq:basis-interpolating}
&& \text{Interpolating:} \qquad && \lagrangeRef[i] (\quadNode[i]) &= \delta_{ij}\\
\label{eq:basis-orthogonality}
&& \text{Orthogonality:} \qquad &&
\underbrace{\int_0^1 \cdots \int_0^1}_{d \,\text{times}}
\lagrangeRef[i] \left( \hat{\bm{x}} \right)
\lagrangeRef[j] \left( \hat{\bm{x}} \right) \dd{\hat{\bm{x}}}
&= \quadWeight[i] \delta_{ij}
    \end{alignat}

We will later need basis functions over a space-cell
$v = 0, \ldots, \NVar, n = 0, \ldots, (N+1)^{\dimensions} - 1, l=0,\ldots,N$
\begin{align}\label{eq:space-basis}
  \begin{split}
  \sbasisRef[][n] (\hat{\bm{x}}) &= \lagrange[n] (x)\\
  \sbasis[][n] (\bm{x}) &= \sbasisRef[][n] \left( \mapping(\bm{x}) \right)
  \end{split}
\end{align}
where $\sbasisRef$ denotes the basis function on the reference cell.
Similarly, we define a basis for a space-time element defined over $(0, 1)^D \times (t^n, t^{n+1})$
\begin{align}\label{eq:space-time-basis}
  \begin{split}
  \stbasisRef[][l,n](\hat{\bm{x}}, \hat{t}) &= \lagrange[l](t) \sbasis[][n] (x) = \lagrange[l](t) \lagrange[n](x)\\
  \stbasis[][l,n] (\bm{x}, t) &= \stbasisRef \left( \mapping(\bm{x}), \timeMapping(t) \right)
  \end{split}
\end{align}
Outside of elements, we define the basis functions to be equal to 0.
\todo{Bold x, bold t?}
\todo{Variables in basis, tensor product?}

\sidetitle{Predictor}
We are now ready to derive the predictor.
Let $t^n$ be the current simulation time and $t^{n+1} = t^n + \Delta t$ the time at the end of a timestep\todo{Explain in basis functions sec}.
We first multiply the conservation law by a test function of the same function space as the basis, integrate over a space-time-element $\cell$ and replace the vector of conservative variables $\Q$ with the space-time predictor $\stpredictor[\cell]$
\begin{equation}\label{eq:weak-pde-space-time}
\intdt{\intdcell{
    \sttestfunction[\cell](\bm{x}, t)
    \pdv{\stpredictor}{t}
}}
+
\intdt{\intdcell{
    \sttestfunction[\cell](\bm{x}, t)
    \left(
      \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
    \right)
}}
=
\intdt{\intdcell{
  \sttestfunction[\cell] \source(\stpredictor)
}}.
\end{equation}

We integrate the first term by parts in time and the flux divergence in space.
Here we do not use the Riemann solver for the flux boundary term but rather use the discrete solution at time $t$\todo{Insert in equation!}.
This corresponds to upwinding in time~\cite{dumbser2008unified}.
Note that this neglects the interaction with neighbouring cells.
We account for this in the corrector step.
\begin{align}\label{eq:space-time-predictor}
\begin{split}
\intdcell{
  \sttestfunction[\cell][k] (\bm{x}, t^{n+1}) \stpredictor(\bm{x}, t^{n+1})
}
&-
\intdt{\intdcell{
    \pdv{}{t} \sttestfunction[\cell][k] (\bm{x}, t) \stpredictor(\bm{x}, t)
}}
= \\
\intdcell{
  \sttestfunction[\cell][k] (\bm{x},t^n) \stpredictor(\bm{x}, t^n)
}
&+
\intdt{\intdcell{
    \sttestfunction[\cell][k] (\bm{x},t) \divergence{\flux(\stpredictor, \gradient{\stpredictor})}
}}
+
\intdt{\intdcell{
    \sttestfunction[\cell][k] (\bm{x}, t) \source(\stpredictor)
}}
\end{split}
\end{align}
\NewDocumentCommand{\stpredictorCoeff}{ O{} O{} }{\underline{\overline{q}}^{#1}_{#2}}
\NewDocumentCommand{\stflux}{ O{} O{} }{\overline{F}^{#1}_{#2}}
\NewDocumentCommand{\stfluxCoeff}{ O{} O{} }{\underline{\overline{F}}^{#1}_{#2}}
\NewDocumentCommand{\stsource}{ O{} O{} }{\overline{S}^{#1}_{#2}}
\NewDocumentCommand{\stsourceCoeff}{ O{} O{} }{\underline{\overline{S}}^{#1}_{#2}}
\NewDocumentCommand{\spred}{ O{} O{} }{u^{#1}_{#2}}
\NewDocumentCommand{\ssolCoeff}{ O{} O{} }{\underline{u}^{#1}_{#2}}
\NewDocumentCommand{\sflux}{ O{} O{} }{F^{#1}_{#2}}
\NewDocumentCommand{\sfluxCoeff}{ O{} O{} }{\underline{F}^{#1}_{#2}}
\NewDocumentCommand{\ssource}{ O{} O{} }{S^{#1}_{#2}}
\NewDocumentCommand{\ssourceCoeff}{ O{} O{} }{\underline{S}^{#1}_{#2}}
All quantities can be expanded in the space-time basis.
For example, the expansion of space-time predictor can be written as
\begin{equation}
  \label{eq:cell-approx-space-time}
  \stpredictor[\cell] (\bm{x}, t) = \stpredictorCoeff[\cell][v,l,n] \stbasis[\cell][v,l,n](\bm{x}, t).
\end{equation}
We denote the space-time flux as $\stflux$ and the space-time source as $\stsource$ with coefficient vectors $\stfluxCoeff$ and $\stsourceCoeff$.
This is a part where we notice the computational advantage of the nodal approach:
The coefficients correspond to the quantity evaluated at the basis node!
